---
title: "Project 1"
author: "Chloe Chen & Kangrui Liu"
date: "9/7/2023"
output:  pdf_document
number_sections: yes
fontsize: 12pt
---
```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE,
                      fig.width=4.5, fig.height=3.5)
```
# 1. Use \texttt{hprice} in \texttt{faraway} package.
```{r}
# prepare data set
library(faraway)
data(hprice)
hprice$homeprice <- exp(hprice$narsp)*1000
```


## 1) What are the mean and the variance of \texttt{homeprice}? What do they mean?
```{r}
meanHprice <- mean(hprice$homeprice)
varHprice <- var(hprice$homeprice)
```
* a) The mean is $94411.42$ and the variance is $1583110349$.\
* b) The terms "mean" and "variance" refer to the average of a data set and the deviation of a data point from the mean, respectively.\

## 2) Construct a 95% confidence interval of the average homeprice. What does the confidence interval imply?
```{r}
n <- dim(hprice)[[1]] # find the number of homeprice
samvar <- var(hprice$homeprice)/(n-1) # sampling var 
t.score <- qt(p=.05/2, df=n-1, lower.tail=F) # 
lowCI <- meanHprice - t.score * sqrt(varHprice)
upCI <- meanHprice + t.score * sqrt(varHprice)
t.test(hprice$homeprice, conf.level = 0.95)
```
* The 95% confidence interval of $\mu_{homeproce}$ is $[90062.70, 98760.14]$\

## 3) Estimate the average homeprice by whether the MSA was adjacent to a coastline, noted in ajwtr, and the standard errors.
```{r}
subset1 <- subset(hprice, hprice$ajwtr == 1)# yes
subset2 <- subset(hprice, hprice$ajwtr == 0)# no
mean1 <- mean(subset1$homeprice)
mean2 <- mean(subset2$homeprice)
#
sum_of_squares_1 <- 0
for (i in 1:nrow(subset1)) {
  sum_of_squares_1 <- sum_of_squares_1 + (subset1$homeprice[i] - mean1)^2
}
s2_1 <- sqrt(sum_of_squares_1 / (nrow(subset1) - 1))
sum_of_squares_2 <- 0
for (i in 1:nrow(subset2)) {
  sum_of_squares_2 <- sum_of_squares_2 + (subset2$homeprice[i] - mean2)^2
}
s2_2 <- sqrt(sum_of_squares_2 / (nrow(subset2) - 1))
```

* The formula of sample variance is $s^2=\frac{\sum_{i=1}^{n}{(y_{i}-\hat{\bar{y}})}^2}{n-1}$.\

## 4) Test the difference in homeprice between coastline MSAs and non-coastline MSAs. Clearly state the formula for the hypothesis, the test method and your rationale for selecting the method. What do you conclude about the hypothesis?
* Step1: Check means and variances \
```{r}
library(dplyr)
hprice%>%
  group_by(ajwtr)%>%
  summarize(m=mean(homeprice),
            v=var(homeprice))
```
* $\mu_{ajwtr=0}=\hat{\bar{y}}_{ajwtr=0}=82388.89$ and $\sigma^2_{ajwtr=0}=s^2_{ajwtr=0}=285315565$
* $\mu_{ajwtr=1}=\hat{\bar{y}}_{ajwtr=1}=111242.96$ and $\sigma^2_{ajwtr=1}=s^2_{ajwtr=1}=2926428411$
* Step2: Test equal variance \
* Testing Equal Variance with F test.
* Corresponding hypothesis: $H_0:\sigma^2_{ajwtr=0}=\sigma^2_{ajwtr=1}$ vs. $H_A:\sigma^2_{ajwtr=0}\neq\sigma^2_{ajwtr=1}$.\
```{r}
var.test(homeprice ~ ajwtr, hprice, alternative = "two.sided")
```
* The range of 95 percent confidence interval is $[0.07088604,0.13297389]$, and the ratio of variances is $0.09749617$. Therefore, We conclude that $H_0:\sigma^2_{ajwtr=0}=\sigma^2_{ajwtr=1}$ does not hold.
* Step3: Conduct proper testing \
* Corresponding hypothesis: $H_0:\mu_{ajwtr=0}=\mu_{ajwtr=1}$ vs. $H_A:\mu_{ajwtr=0}\neq\mu_{ajwtr=1}$.\
```{r}
t.test(homeprice ~ ajwtr, hprice, var.equal=FALSE, conf.int = 0.95)
```
* We reject $H_0:\mu_{ajwtr=0}=\mu_{ajwtr=1}$.\

## 5) Estimate the Pearson correlation coefficient between homeprice and per capita income of the MSA of a given year, noted in ypc.
```{r}
cor(hprice$homeprice,hprice$ypc,method = "pearson")
```
* The Pearson correlation coefficient between homeprice and ypc is $0.7437474$.

## 6) Test whether the correlation coefficient between homeprice and ypc is 0 or not. Clearly state the hypothesis including the formula. What do you conclude?
```{r}
cor.test(hprice$homeprice,hprice$ypc,method = "pearson")
```
* Corresponding hypothesis: $H_0:\rho=0$ vs. $H_A:\rho\neq0$.
* $\bar\rho=0.7437474$, and 95% CI is $[0.6907661,0.7887854]$, So reject $H_0$.

## 7) Can you say that per capita income has an effect on the home sales price using the results from #6)? Why or why not?
```{r}
cor.test(hprice$ypc,hprice$homeprice,method = "pearson")
```
* Yes, they both are related to other. Because correlation...

## 8) Test the normality of homeprice. Would this test result change your responses to #1) to 7)? Why or why not?
```{r}
library(ggplot2)
ggplot(hprice, aes(x=homeprice)) + geom_histogram()
```
```{r, error=TRUE}
qqnorm(hprice$homeprice, main="hprice", ylab="y_{i:n}", xlab="m_{i:n}")+
qqline(hprice$homeprice, col="red",lwd=2)
```
```{r}
shapiro.test(hprice$homeprice)
```
* In Shapiro-Wilk normality test, p-value < 0.05, and from the histogram, as well as the qq-plot, homeprice does not appear to follow a normal distribution.
