---
title: "Project 1"
author: "Chloe Chen & Kangrui Liu"
date: "9/5/2023"
output:  pdf_document
number_sections: yes
fontsize: 12pt
---
# 1. Use \texttt{hprice} in \texttt{faraway} package.

## 1) What are the mean and the variance of homeprice? What do they mean?
```{r}

```

## 2) Construct a 95% confidence interval of the average homeprice. What does the confidence interval imply?
```{r}

```

## 3) Estimate the average homeprice by whether the MAS was adjacent to a coastline, noted in ajwtr, and the standard errors.
```{r}

```

## 4) Test the difference in homeprice between coastline MSAs and non-coastline MSAs. Clearly state the formula for the hypothesis, the test method and your rationale for selecting the method. What do you conclude about the hypothesis?
```{r}

```

## 5) Estimate the Pearson correlation coefficient between homeprice and per capita income of the MSA of a given year, noted in ypc.
```{r}

```

## 6) Test whether the correlation coefficient between homeprice and ypc is 0 or not. Clearly state the hypothesis including the formula. What do you conclude?
```{r}

```

## 7) Can you say that per capita income has an effect on the home sales price using the results from #6)? Why or why not?
```{r}

```

## 8) Test the normality of homeprice. Would this test result change your responses to #1) to 7)? Why or why not?
```{r}

```

