---
title: "Project4"
author: "Chloe Chen & Kangrui Liu"
date: "2023-09-26"
output: pdf_document
number_sections: yes
fontsize: 12pt
---

```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE,
                      fig.width=4.5, fig.height=3)
```

This exercise involves the Boston housing data set in ISLR2. Assume that we are
interested in median home values, medv.
```{r}
library(ISLR2)
data(Boston)
```

## 1. Examine *medv* as a function of *crim*, *zn* and *indus* in a multiple linear regression.
### A.Identify predictors significant at 5% level.
```{r}
m<-lm(medv~crim+zn+indus,Boston)
summary(m)
```

### B.List the hypotheses tested in 1.A and their conclusions.
* $H_0$:${\beta}_0=0$; $H_1$:${\beta}_0\neq 0$
* $H_0$:${\beta}_{crim}=0$; $H_1$:${\beta}_{crim}\neq 0$
* $H_0$:${\beta}_{zn}=0$; $H_1$:${\beta}_{zn}\neq 0$
* $H_0$:${\beta}_{indus}=0$; $H_1$:${\beta}_{indus}\neq 0$
- The p-values for all of the coefficients are lower than 0.05, so we reject $H_0$, they are all significantly not equal to 0.

### C.What do the estimated regression coefficients in 1.A mean in lay terms? Do
they make sense?
- The intercept:The expected mean of median value of owner-occupied homes(*medv*) when all the predictors are $0$. This is almost impossible to happen in real life, so it does not make sense.
- ${\beta}_{crim}$:the expected change in median value of owner-occupied homes(*medv*) is $-0.24863$ for every one unit change in per capita crime rate(*crim*) while holding the other predictors constant.
- ${\beta}_{zn}$:the expected change in median value of owner-occupied homes(*medv*) is $0.05850$ for every one unit change in proportion of residential land zoned for lots(*zn*) while holding the other predictors constant.
- ${\beta}_{indus}$:the expected change in median value of owner-occupied homes(*medv*) is $-0.41558$ for every one unit change in proportion of non-retail business acres per town(*indus*) while holding the other predictors constant.

### D.Construct 95% confidence intervals for ${\beta}_crim$, ${\beta}_zn$ and ${\beta}_indus$. How do the confidence intervals correspond to 1.A and 1.B?
```{r}
CI_crim <- confint(m, "crim", level = 0.95)
CI_zn <- confint(m, "zn", level = 0.95)
CI_indus <- confint(m, "indus", level = 0.95)
print("The 95% confidence interval for crim:")
print(CI_crim)

print("The 95% confidence interval for zn:")
print(CI_zn)

print("The 95% confidence interval for indus:")
print(CI_indus)

```
- Based on the confidence interval values and the coefficients of each predictor variable, we find that the coefficients are indeed significant at the 95% level and are able to reject $H_0$.

### E.Calculate the ${R^2}$ and $R^2_{adj}$ by hand and report whether this matches the ${R^2}$ and $R^2_{adj}$ from the output and whey mean?
```{r}


```

2. Compare the model from #1 and a simple linear regression of *medv* as a
function of *zn*. Which would you prefer?
```{r}
m_zn<-lm(medv~zn,Boston)
summary(m_zn)
summary(m)
```

