---
title: "Project4"
author: "Chloe Chen & Kangrui Liu"
date: "2023-09-26"
output: pdf_document
number_sections: yes
fontsize: 12pt
---

```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE,
                      fig.width=4.5, fig.height=3)
```

This exercise involves the Boston housing data set in ISLR2. Assume that we are
interested in median home values, medv.
```{r}
library(ISLR2)
data(Boston)
```

## 1. Examine *medv* as a function of *crim*, *zn* and *indus* in a multiple linear regression.
### A.Identify predictors significant at 5% level.
```{r}
m<-lm(medv~crim+zn+indus,Boston)
summary(m)
```
- the p-value for *crim* is $2.52e-08$; the p-value for *zn* is $0.000889$; the p-value for *indus* is $1.77e-10$.

### B.List the hypotheses tested in 1.A and their conclusions.
* $H_0$:${\beta}_0=0$; $H_1$:${\beta}_0\neq 0$
* $H_0$:${\beta}_{crim}=0$; $H_1$:${\beta}_{crim}\neq 0$
* $H_0$:${\beta}_{zn}=0$; $H_1$:${\beta}_{zn}\neq 0$
* $H_0$:${\beta}_{indus}=0$; $H_1$:${\beta}_{indus}\neq 0$
- Conclusions: The p-values for all of the coefficients are lower than 0.05, so we reject $H_0$, they are all significantly not equal to 0, that means all of the coefficients of the predictors are statistically significant.

### C.What do the estimated regression coefficients in 1.A mean in lay terms? Do
they make sense?
- The intercept($\beta_0=27.39465$):The expected mean of median value of owner-occupied homes(*medv*) is $27.39465$ when all the predictors are $0$. This is almost impossible to happen in real life, so it does not make sense.
- ${\beta}_{crim}=-0.24863$:the expected change in median value of owner-occupied homes(*medv*) is $-0.24863$ for every one unit change in per capita crime rate(*crim*) while holding the other predictors constant.
- ${\beta}_{zn}=0.05850$:the expected change in median value of owner-occupied homes(*medv*) is $0.05850$ for every one unit change in proportion of residential land zoned for lots(*zn*) while holding the other predictors constant.
- ${\beta}_{indus}-0.41558$:the expected change in median value of owner-occupied homes(*medv*) is $-0.41558$ for every one unit change in proportion of non-retail business acres per town(*indus*) while holding the other predictors constant.

### D.Construct 95% confidence intervals for ${\beta}_crim$, ${\beta}_zn$ and ${\beta}_indus$. How do the confidence intervals correspond to 1.A and 1.B?
```{r}
CI <- confint(model)
CI

```
- $\beta_0$ intercept falls between 25.70 and 29.09
- $\beta_1$ crim falls between -0.33 and -0.16
- $\beta_2$ zn falls between 0.02 and 0.093
- $\beta_3$ indus falls between -0.54 and -0.29
### Based on the confidence interval values and the coefficients of each predictor variable, we find that the coefficients are indeed significant at the 95% level and are able to reject $H_0$.In lay terms, we have 95% confidence to say that the predictors are not equal to 0.

### E.Calculate the ${R^2}$ and $R^2_{adj}$ by hand and report whether this matches the ${R^2}$ and $R^2_{adj}$ from the output and whey mean?
```{r}
# Given regression coefficients
intercept <- 27.39465
coef_crim <- -0.24863
coef_zn <- 0.05850
coef_indus <- -0.41558
k<- 3
n<- dim(Boston)[1]

# Predicted values
predicted_m <- intercept + coef_crim * Boston$crim + coef_zn * Boston$zn + coef_indus * Boston$indus
# SSR (sum of squared residuals)
SSR <- sum((predicted_m - Boston$medv)^2)
# SST (total sum of squares)
mean_medv <- mean(Boston$medv)
SST <- sum((Boston$medv - mean_medv)^2)
# Calculate R-squared & adjusted_R_squared
R_squared <- 1 - (SSR / SST)
adjusted_R_squared <- 1 - ((1 - R_squared) * (n - 1)) / (n - k - 1)

R_squared; adjusted_R_squared

```
- The value of ${R^2}$ is $0.2937$ ; the value of $R^2_{adj}$ is $0.2895$ .This matches the ${R^2}$ and $R^2_{adj}$ from the previous output.
- R squared means the proportion of data is explained by model. Adjusted R squared takes DF(the number of predictors) into account considering parsimony. 

2. Compare the model from #1 and a simple linear regression of *medv* as a
function of *zn*. Which would you prefer?
```{r}
m_zn<-lm(medv~zn,Boston)
summary(m_zn)
summary(m)
```

- Theoretically, we would prefer the model from #1 based on the higher adjusted $R^2$ value. 
- Nonetheless, the final choice should be guided by a comprehensive evaluation of the models and their suitability for specific research or analytical goals.In addition to adjusted $R^2$, it is also crucial to take into account additional aspects, such as the theoretical or practical significance of the predictors, domain knowledge, and the specific goals of certain study.We are also obliged to further evaluate the regression models' underlying assumptions and carry out diagnostic tests.
