---
title: "Proj 8"
author: "Chloe Chen & Kangrui Liu"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
number_sections: yes
fontsize: 12pt
---


```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo=TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE,
                      fig.width=4.5, fig.height=3)
```



## 1. Faraway Chapter 7. Exercise 5.
* For the prostate data, fit a model with lpsa as the response and the other variables
as predictors.
```{r}
library(faraway)
data(prostate)
mod1<-lm(lpsa~lcavol+lweight+age+lbph+svi+lcp+gleason+pgg45, prostate)
summary(mod1)
```
### 1.A. Compute and comment on the condition numbers.
```{r}
x_mod1<-model.matrix(mod1)[,-1]
x_mod1
e1<-eigen(t(x_mod1)%*%x_mod1) # * is multiply in R %*% is matrix multiplication
e1
e1$val
kappa1<-sqrt(max(e1$val)/min(e1$val))
nu1<-sqrt(max(e1$val)/e1$val)
kappa1;nu1 
```
* The largest condition number(243.30248) is equal to kappa, and is >30.

### 1.B. Compute and comment on the correlations between the predictors.
```{r}
cor(prostate[,c("lcavol","lweight","age","lbph","svi","lcp","gleason","pgg45")])
```
* Most of the variables are having a weak linear relationship.
* The correlations between lcp and lvocal, svi, pgg45 are kind of large, the correlation betweengleason and pgg45 is kind of large， which would imply they are related to some extent.

### 1.C. Compute the variance inflation factors.

```{r}
vif(mod1)
```
* Most of the variables have relatively low VIF, for example, lweight, age, and lbph, the VIF are close to 1, so there's no strong reason to believe they are having a significant multicollinearity issue.


## 2. Faraway Chapter 8. Exercise 4.
* For the cars dataset, fit a linear model with distance as the response and speed as the predictor.
```{r}
data(cars)
mod2<-lm(dist~speed, cars)
summary(mod2)
```


### 2.A. Test the homoscedasticity assumption using both a scatter plot between the residuals and fitted values and an F-test of equal variance below and above the fitted value of 30.
```{r}
plot(fitted(mod2),residuals(mod2), xlab = "Fitted Values", ylab = "Residuals")+
abline(h=0,col="red")
res_mod2<-residuals(mod2)
fitted_values <- fitted(mod2)
#below and above thefitted value of 30
below <- res_mod2[fitted_values < 30]
above <- res_mod2[fitted_values >= 30]
var_test<- var.test(below,above)
var_test
```
* From the p-value we can reject the null hypothesis of equal variances between the two groups.

### 2.B. Report the estimate of the heteroscedastic consistent variance for the
regression slope.
```{r}
library(dplyr)
library(sandwich)
library(lmtest)

hetvar <- mod2 %>% 
  vcovHC() %>% # calculate Heteroscedasticity-consistent estimation of the covariance matrix for coefficients
  diag() %>% # gives variances as they are the diagonal of the covariance matrix
  sqrt() # gives standard error as it is square root of variance
summary(mod2)
speed_htvar <- hetvar["speed"]
speed_htvar

```
* The heteroscedastic consistent variance for the regression slope is $0.4275372$.


### 2.C. Construct 95% confidence interval of the regression slope assuming homoscedasticity and using the results in 2.B. How do they compare?
```{r}
# Calculate 95% CI using heteroscedasticity-consistent variance
mod2$coefficients[2]+c(-1,1)*qt(0.975,mod2$df.residual)*sqrt(hetvar[2])
```
* The results in 2.B. lies in the 95% confidence interval(wider than homoscedasticity-consistent interval).
* This is because after considering the heteroscedasticity, the uncertainty will be higher which means standard error would be larger.


### 2.D. Check for the lack of fit of the model.
```{r}
mod2_RESID<-lm(resid(mod2)~factor(speed),cars)
anova(mod2)
anova(mod2_RESID)

length(unique(cars$speed))
length(cars$speed)
(((11354 - 6764.8)/(48-31))/(6764.8/31))

1 - pf(1.237071,df1=31,df2=17)

library(olsrr)
ols_pure_error_anova(mod2)

```
* We can see from the Lack of Fit F Test that the F value is 1.23695, the p value is 0.295 (>.05) for the "lack of test". So there's no sufficient reason to believe the model is lack of fit.